Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-07-04 WSSSPE2_report/WSSSPE2_review.tex
undoing some changes

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Papers}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
The papers that were discussed in the Exploring Sustainability group are:
\begin{itemize}
\item Mario Rosado de Souza, Robert Haines, and Caroline Jay. Defining
sustainability through developers' eyes: Recommendations from an interview
study~\cite{wssspe2_rosada_de_souza}

\item Robert Downs, W. Christopher Lenhardt, Erin Robinson, Ethan Davis, and
Nicholas Weber. Community recommendations for sustainable scientific
software~\cite{wssspe2_downs}

\item Abani Patra, Matthew Jones, Steven Gallo, Kyle Marcus, and Tevfik Kosar.
Role of online platforms, communications and workflows in developing sustainable
software for science communities~\cite{wssspe2_patra}

\item Marlon Pierce, Suresh Marru, and Chris Mattmann. {WSSSPE2}: Patching it up,
pulling it forward~\cite{wssspe2_pierce}

\item Justin Shi. Seeking the principles of sustainable software
engineering~\cite{wssspe2_shi}

\item Colin C. Venters, Michael K. Griffiths, Violeta Holmes, Rupert R. Ward, and
David J. Cooke. The Nebuchadnezzar effect: dreaming of sustainable software
through sustainable software architectures~\cite{wssspe2_venters}
\end{itemize}

@@

Group discussion started with a position statement by each person surrounding
what they had learned from their software development experiences and how those
lessons might be translated into actionable outcomes. Participants came from a
range of backgrounds, and represented multiple software and training
initiatives, including Software Carpentry (SwC), Data Carpentry (DC), Open Science for
Synthesis (OSS), the Community Surface Dynamics Modeling Systems group, ROpenSci,
DataONE, the HDF Group, and others. Some of these software experiences were
focused on development of new products for use in the sciences (e.g., ROpenSci,
CSDMS), and these recounted the difficulties of engaging with disciplinary
scientists in writing software. Software was clearly utilized broadly across the
various science disciplines represented, and it was developed within those
disciplines as well. Some researchers created software for statistical analysis
and modeling, while others used it to control instrumentation, collate data
across networks, collect information from respondents, and many other uses.
There was broad consensus that, within the disciplines represented, formal
training in any type of software development or engineering was rare among the
practitioners; most are self-taught, and develop software to get another job
done. Any ancillary utility of the software outside of the specific science
target was generally unplanned and few researchers would want to invest more
time to make their own software more re-usable.

@@

Three main actions were identified by Subgroup A that would be of interest to
participants and benefit the research community. These focused on the
desire to amplify the current community efforts in training and software
engineering by connecting the current training initiatives (e.g., SwC, DC, 
rOpenSci, OSS) that are of different durations at appropriate stages.
Generally it was felt that a modicum of interaction between short term workshops
(SwC, DC), medium term trainings (e.g., OSS), and longer-term courses (e.g.,
BIDS Data Science) would benefit from coordinating curriculum, discussing and
aligning prerequisites, and coordinating timing of courses. A training
coordination effort would go a long way towards amplifying the value of the
individual efforts and make them all more effective.

>>>>

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Papers}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
The papers that were discussed in the Exploring Sustainability group are:
\begin{itemize}
\item Mario Rosado de Souza, Robert Haines, and Caroline Jay. Defining
Sustainability through Developers' Eyes: Recommendations from an Interview
Study~\cite{wssspe2_rosada_de_souza}

\item Robert Downs, W. Christopher Lenhardt, Erin Robinson, Ethan Davis, and
Nicholas Weber. Community Recommendations for Sustainable Scientific
Software~\cite{wssspe2_downs}

\item Abani Patra, Matthew Jones, Steven Gallo, Kyle Marcus, and Tevfik Kosar.
Role of Online Platforms, Communications and Workflows in Developing Sustainable
Software for Science Communities~\cite{wssspe2_patra}

\item Marlon Pierce, Suresh Marru, and Chris Mattmann. {WSSSPE2}: Patching It Up,
Pulling It Forward~\cite{wssspe2_pierce}

\item Justin Shi. Seeking the Principles of Sustainable Software
Engineering~\cite{wssspe2_shi}

\item Colin C. Venters, Michael K. Griffiths, Violeta Holmes, Rupert R. Ward, and
David J. Cooke. The Nebuchadnezzar Effect: Dreaming of Sustainable Software
through Sustainable Software Architectures~\cite{wssspe2_venters}
\end{itemize}

@@

Group discussion started with a position statement by each person surrounding
what they had learned from their software development experiences and how those
lessons might be translated into actionable outcomes. Participants came from a
range of backgrounds, and represented multiple software and training
initiatives, including Software Carpentry, Data Carpentry, Open Science for
Synthesis, the Community Surface Dynamics Modeling Systems (CSDMS) group, ROpenSci,
DataONE, the HDF Group, and others. Some of these software experiences were
focused on development of new products for use in the sciences (e.g., ROpenSci,
CSDMS), and these recounted the difficulties of engaging with disciplinary
scientists in writing software. Software was clearly utilized broadly across the
various science disciplines represented, and it was developed within those
disciplines as well. Some researchers created software for statistical analysis
and modeling, while others used it to control instrumentation, collate data
across networks, collect information from respondents, and many other uses.
There was broad consensus that, within the disciplines represented, formal
training in any type of software development or engineering was rare among the
practitioners; most are self-taught, and develop software to get another job
done. Any ancillary utility of the software outside of the specific science
target was generally unplanned and few researchers would want to invest more
time to make their own software more re-usable.

@@

Three main actions were identified by Subgroup A that would be of interest to
participants and would benefit the research community. These focused on the
desire to amplify the current community efforts in training and software
engineering by connecting the current training initiatives (e.g., Software and Data Carpentry, 
rOpenSci, OSS) that are of different durations at appropriate stages.
Generally it was felt that a modicum of interaction between short term workshops
(SwC, DC), medium term trainings (e.g., OSS), and longer-term courses (e.g.,
BIDS Data Science) would benefit from coordinating curriculum, discussing and
aligning prerequisites, and coordinating timing of courses. A training
coordination effort would go a long way towards amplifying the value of the
individual efforts and make them all more effective.
